Layered Formal Verification of an Industrial TCP/IP Stack

### 2021-06-01 SecDev21/paper.tex
fixing the synchronous exchanges section

The concurrent implementation of the TCP protocol allows for multiple interactions between the tasks, synchronous (with a call to \spark{TCP_Wait_For_Events}) or asynchronous (between two user function calls), and thus, numerous possible state changes at any given moment. This makes the functional-specification verification significantly challenging, as SPARK does not have a native mode to deal precisely with interactions related to concurrency. To address this, we introduced SPARK functions with contracts that model how the different possible interactions can modify the state of a socket. Finding a correct and precise contract for \spark{TCP_Wait_For_Events} is challenging but needed for the verification of the TCP user functions. The solution will be elaborated in the next section.

@@

Sometimes the resulting state of a user function can be affected by the rest of the TCP tasks is when the user's functions release the mutex to allow for other events to take place. In this case, the function \spark{TCP_Wait_For_Events} is called from the user function to check if the event requested is completed. For example, the \spark{TCP_Connect} user function calls \spark{TCP_Wait_For_Events} to wait for the completion of the socket connection event, namely the \spark{SOCKET\_EVENT\_CONNECTED}. The call to the \spark{TCP_Wait_For_Events} releases the mutex, and when the \emph{arriving} task receives the appropriate segments, in this case, a segment with the \syn flag, the connection will be established. This will move the socket to the \sestab state. In the meantime, the \spark{TCP\_Update\_Events} function is monitoring for any state changes and when it notices that the connection is established it will trigger the \spark{SOCKET\_EVENT\_CONNECTED} event, using the OS event mechanism. This will allow the \spark{TCP\_Connect} user function to resume execution.

>>>>

The concurrent implementation of the TCP protocol allows for multiple interactions between the tasks, synchronous (with a call to \spark{TCP_Wait_For_Events}) or asynchronous (between two user function calls), and thus, numerous possible state changes at any given moment. This makes the functional specification verification significantly challenging, as SPARK does not have a native mode to deal precisely with interactions related to concurrency. To address this, we introduced SPARK functions with contracts that model how the different possible interactions can modify the state of a socket. Finding a correct and precise contract for \spark{TCP_Wait_For_Events} is challenging but needed for the verification of the TCP user functions. The solution is elaborated in the following section.

@@

In the synchronous scenario of interaction between the TCP tasks, the resulting state of a user function can be affected by the rest of the TCP tasks when the user's functions release the mutex to allow for other events to occur. In this case, the function \spark{TCP_Wait_For_Events} is called from the user function to check if the event requested is completed. For example, the \spark{TCP_Connect} user function calls \spark{TCP_Wait_For_Events} to wait for the completion of the socket connection event, namely the \spark{SOCKET\_EVENT\_CONNECTED}. The call to the \spark{TCP_Wait_For_Events} releases the mutex, and when the \emph{arriving} task receives the appropriate segments, in this case, a segment with the \syn flag, the connection will be established. This will move the socket to the \sestab state. In the meantime, the \spark{TCP\_Update\_Events} function is monitoring for any state changes, and when it notices that the connection is established, it will trigger the \spark{SOCKET\_EVENT\_CONNECTED} event, using the OS event mechanism. This will allow the \spark{TCP\_Connect} user function to resume execution.
